```{r include=FALSE}
knitr::knit_child("04-data_manipulation.Rmd")
source("./dependencies.R")
```

# Visualization

## Bar chart
A bar chart is a representation of numerical data in pictorial form of rectangles (or bars) having uniform width and varying heights." They are also known as bar graphs.

```{r}
#construction of the dataframe
data_barchart <- as.data.frame(table(data_1$`What is your religion?`))
data_barchart <- data_barchart %>%
    dplyr::mutate(percentage = round(100*(Freq/sum(Freq)),2),
                  pct1 = paste0(percentage, "%")) %>%
  rename(Religion=Var1)


#plot the bar chart
plotly::plot_ly(data_barchart, x = ~Religion,
                  type = "bar",
                  y = ~percentage,
                  marker = list(color = "#318CE7"),
                  text = paste(data_barchart$pct1, sep = ""), textposition = 'outside') %>%
    layout(title = "Number of persons by religion"
    )

```


## Pie chart
A pie chart is a type of graph representing data in a circular form, with each slice of the circle representing a fraction or proportionate part of the whole.

```{r}
#construction of the dataframe
data_piechart <- as.data.frame(table(data_1$Sex))
data_piechart <- data_piechart %>%
    dplyr::mutate(percentage = round(100*(Freq/sum(Freq)),2),
                  pct1 = paste0(percentage, "%"))

#plot the pie chart
plotly::plot_ly(data_piechart, labels= ~Var1,
          values= ~Freq, type="pie",
          hoverinfo = 'text',
          textinfo = 'label+percent',
          insidetextfont = list(color = '#FFFFFF'),
          text = ~paste("Sex :",Var1,
                        "<br>Number of persons :", Freq,
                        "<br>Percentage :", pct1),
          marker = list(colors = c("#318CE7", "#89CFF0"),
                        line = list(color = '#FFFFFF', width = 1),showlegend = FALSE)) %>%
    layout(title="",
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
```


## Histogram
A histogram is a chart that plots the distribution of a numeric variable’s values as a series of bars. Each bar typically covers a range of numeric values called a bin or class; a bar’s height indicates the frequency of data points with a value within the corresponding bin.
```{r}
library(ggplot2)

# Change colors
p<-ggplot(data_1, aes(x=`How many children do you have?`)) + 
  geom_histogram(color="black", fill="white")
p
```


## Scatter plot

## Line chart

## Map visualization

## Wordcloud